Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-10-08 paper.tex
Almost ready.
Need to read it one more time

\author{Ryan P. McDonnell} \email{rpmcdonnell@wisc.edu}
\author{Daniel D. Kohler}
\author{John C. Wright} \email{wright@chem.wisc.edu}

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of nonlinear spectroscopy methods that form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling and correlations between different vibrational, electronic, and vibronic states. \cite{RN281, RN103, Cho2008} 
Four wave mixing CMDS has resolved anharmonicities, ultrafast dynamics, and other inter- and intra-molecular couplings in numerous systems. \cite{Cho2008, Gaynor2017, Ziegler2018, Ogilvie2019, Bonn2021, RN325, Chen2024}
To target vibrational and electronic coupling, methods commonly make use of Raman transitions, e.g., doubly vibrationally enhanced spectroscopy (DOVE), triple sum frequency (TSF), coherent anti-Stokes Raman spectroscopy (CARS), resonant impulse stimulated Raman spectroscopy (RISRS).\cite{RN103, Dhar1994, RN345, Cho2000, RN491}
Here we explore the less-utilized hyper-Raman transition in four wave mixing CMDS.

@@

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a$ and $\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$. 
While most hyper-Raman experiments are performed using degenerate input pulses ($\omega_a = \omega_b$), it is possible to perform the technique with non-degenerate input pulses ($\omega_a \neq \omega_b$). \cite{Denisov1986, Kozich2007}
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through a two-photon inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been observed in several CMDS studies (sometimes under the moniker of SIVE).\cite{Zilian1994, RN350, RN351, RN352, RN353, Chen1998, RN362, RN418, Wang2021, Bonn2024, McDonnell2024}

@@

Nonlinear spectroscopic probes of electronic-vibrational coupling are not limited to the Raman and hyper-Raman based techniques diagrammed in \autoref{fig:comparisonwmel}.
Independently developed by the Fleming and Khalil groups, respectively, 2D Electronic-Vibrational (2D-EV) and 2D Vibrational-Electronic (2D-VE) spectroscopy are based upon pump-probe and absorption type pathways which also investigate e-v coupling.\cite{Oliver2014, Courtney2015, Courtney2015_1}
The Rao group has also developed a $\chi^{(4)}$ analogue of 2D-EV, 2D-EV sum frequency generation (2D-EVSFG), to probe interfacial e-v coupling. \cite{Deng2021}  
These methods have been used to understand vibronic coupling in a variety of molecular systems and, albeit different selection rules, serve as 2D pump-probe type analogues to the infrared-hyper-Raman spectroscopies discussed here.

When the hyper-Raman excitation in HDFG or HSFG is resonant or near-resonant with electronic states, the brightness of the vibrational feature depends on the vibronic structure of the electronic state.
The electronic spectrum probed in HDFG or HSFG can inform on processes that control ultrafast electronic relaxation in molecular and biological systems in ways similar to resonance Raman.\cite{Cho2001, Bredenbeck2015, Arsenault2021}
Since the selection rules of hyper-Raman scattering differ from Raman scattering, the hyper-Raman excitation spectra give a unique alternative to Raman scattering to understand vibrational spectra, electronic structure and vibronic coupling in molecular systems. \cite{Olson2018}

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
Coherent spectroscopies are often described using the density matrix formalism, where an n$^{\text{th}}$ order density matrix $\rho^{(n)}$ is propagated using the Liouville-von Neumann equation:
\begin{equation}\label{LVNE}
	\dot{\rho}^{(n)} = \frac{1}{i \hbar} [H + V(t), \rho^{(n-1)}] + \dot{\rho}^{(n)}_R,
\end{equation}
where $H$ is the (static) system Hamiltonian, $V(t)$ is the field-matter interaction and $\rho_R$ is the relaxation density matrix. \cite{RN455}
In systems where interactions with surroundings are relevant, expressions for $\dot{\rho}_R$ may be quite complicated due to intricate phenomena such as system-bath interactions and solvent-solute coupling, among others. \cite{Sung2001}
In the limit where input field pulsewidths are much longer than electronic dephasing times and the individual molecule limit is approached, it becomes acceptable to write a time-independent relaxation matrix $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$, i.e. exponential relaxation, where $\Gamma$ is the system's dephasing rate. \cite{RN455, Sue1986, Sung2001}
While exponential relaxation is usually an inappropriate description of most physical systems involving solvent-solute interactions,\cite{Sue1986, Yan1988, Li1994, Myers1997} our goal is to introduce simple, quantitative models to understand HDFG response.
Thus, we write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$.
The exponential relaxation limit has proven useful in investigating the properties of 2D-EV, 2D-VE, and doubly resonant SFG using pulses much longer than electronic dephasing times. \cite{Raschke2002, Gaynor2017}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.\cite{Cho2008}
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or obscure coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=\textwidth / 2]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
		The virtual states are arbitrarily placed and fluctuate in energy dependent upon the energy of the input fields.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is vibrationally resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned. \cite{McDonnell2024}
Note that in the case where $\omega_1$ interacts first but $\omega_2$ is resonant, the nonlinear output generated nearly vanishes as the first resonant pulse arrives after a non-resonant pulse.

@@

Other potential interferants in HDFG spectroscopy are direct and sequential cascades of second-order vDFG and vSFG processes.\cite{RN297, Cho2000_Cascade, RN301}
While unimportant in achiral isotropic systems,\cite{Belkin2000} second-order cascades may become important in media where SFG and DFG are allowed, e.g. chiral media, interfaces or non-centrosymmetric media. 
Nevertheless, in the discussion that follows for the rest of this manuscript, we will assume negligible second-order cascades.

It is useful to expose relationships between nonlinear output, transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit ($\dot{\rho}^{(n)} = 0 \ \forall n$). \cite{Moad2005}
Under the electric dipole approximation, the $I^\text{th}$ Cartesian component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation) \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(\omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the $IJKL$ element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}\label{eq:nfgamma}
	\chi^{(3)}_{IJKL} = NF \langle \gamma_{ijkl} \rangle_{IJKL}.
\end{equation}
For clarity, we define the input electric fields as
\begin{equation}
	\vec{E}\left(\vec{x},t\right) = \frac{1}{2} \vec{E}_0 \exp(i\left(\vec{k}\cdot\vec{x} -\omega t\right)) +\text{c.c.}.
\end{equation}
In \autoref{eq:nfgamma}, $N$ is a number density, $F=\prod_j  \frac{1}{3} \left({\left(n(\omega_j)\right)^2 + 1} \right)$ is the Lorentz local field factor, where $n$ is the (frequency dependent) refractive index and $\gamma_{ijkl}$ is the third-order polarizability (i.e., second hyperpolarizability). \cite{Bedeaux1973}
The brackets indicate an orientational average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.
The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}

By propagating \autoref{LVNE} in the steady state limit, the HDFG hyperpolarizability is \cite{Prior1984}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl}^{vg} =	- \sum_{e,m} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{ve}_{i} \mu^{em}_{j} \mu^{mg}_{k} \mu^{gv}_{l} }{\Delta_{ev}^{3+2-1} \Delta_{mv}^{2-1}\Delta_{gv}^{-1}}  \rho_{gg}
\end{equation}
where: $\mu^{ab}_{j}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl}^{-x+y+z} = \omega_{kl} - (-\omega_{x}+\omega_{y} + \omega_{z}) - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the $\rho_{kl}$ dephasing, and $\rho_{gg}$ is the ground state population.
Here, $\ket{e}$ and $\ket{m}$ are virtual states, and thus necessarily summed over in \autoref{sivegamma}.
The pulse identifiers on the $\Delta_{kl}$ terms will be suppressed henceforth for legibility.
$D$, the Maker-Terhune degeneracy factor, accounts for permutation symmetry of the excitation fields.\cite{RN134} 
For a HDFG experiment using two (three) distinct, non-degenerate input fields, $D = 3 (6)$.

@@

It is useful to investigate the selection rules without vibronic resonance to understand how HDFG can be used as an upconverted infrared spectroscopy.
We first investigate the HDFG selection rules through a Placzek type formalism.
When $\omega_2+\omega_3$ is significantly detuned from resonance and the vibronic nature of the virtual states is ignored, the Placzek approximation can be invoked to study selection rules in terms of the hyper-Raman hyperpolarizability, $\beta$. \cite{Placzek1934, Long1970}
Summing over the virtual states $\ket{e}$ and $\ket{m}$ forms the hyper-Raman hyperpolarizability so that\cite{Long1970} 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{vg}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}.
\end{equation}
Since all infrared active transitions are hyper-Raman active, HDFG is allowed for any infrared active transition. \cite{Andrews1978}
This selection rule is generally valid for any HDFG or HSFG process when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

The general selection rule can be simplified by considering the excitation of an infrared active normal mode.  
Assume $\ket{v}$ is the $v^{\text{th}}$ normal mode.
We Taylor expand the dipole and first hyperpolarizability operators to first order in the normal mode coordinate $Q_v$ about equilibrium:
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_v}\right)_0 Q_v 
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_v}\right)_0 Q_v
	\end{equation}
\end{subequations}
Taking $\ket{g} = \ket{0}, \ket{v} = \ket{1}$, substituting into \autoref{sivebeta} and using typical Harmonic Oscillator selection rules, gives the HDFG hyperpolarizability to $\order{Q_v}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{8D \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{vg}_{ijk}}{\partial Q_v}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_v}}\right)_0  \rho_{gg}
\end{equation}
where we have used $\mel{v}{\mu_{l,0}}{g} = 0 = \mel{1}{\beta_{ijk,0}}{0}$, and $\mel{1}{Q_v}{0} = \sqrt{\frac{\hbar}{2\omega_{vg}}}$.
Here, $\omega_{vg}$ is the characteristic $\ket{0} \rightarrow \ket{1}$ frequency.\cite{RN459}
Since \autoref{SIVEselection} is non-zero in the harmonic oscillator limit, HDFG output is allowed for infrared active normal modes in the harmonic limit.

\subsection{Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the general source of HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to Albrecht's treatment of Raman spectroscopy (i.e., sum-over-states formalism).\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(\textbf{Q})) \otimes \ket{b}$ for electronic states $\{|a(\textbf{Q}))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where \textbf{Q}, the system normal modes, is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on \textbf{Q} for simplicity.
It is worthwhile to note that the time-dependent Raman formalism developed by Heller \textit{et al.} can be applied to hyper-Raman scattering. \cite{Lee1979, Silverstein2012}
However, since our goal is to expose the Quantum Mechanical aspects of HDFG spectroscopy, we choose to employ the typical sum-over-states formalism.

>>>>

\author{Ryan P. McDonnell} %\email{rpmcdonnell@wisc.edu}
\author{Daniel D. Kohler}
\author{John C. Wright} \email{wright@chem.wisc.edu}

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of nonlinear spectroscopy methods that form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling and correlations between different vibrational, electronic, and vibronic states. \cite{RN281, RN103, Cho2008} 
Four wave mixing CMDS has resolved anharmonicities, ultrafast dynamics, and other inter- and intra-molecular couplings in numerous systems. \cite{Cho2008, Gaynor2017, Ziegler2018, Ogilvie2019, Bonn2021, RN325, Chen2024}
To target vibrational and electronic coupling, methods commonly make use of Raman transitions, e.g., doubly vibrationally enhanced spectroscopy (DOVE), triple sum frequency (TSF), coherent anti-Stokes Raman spectroscopy (CARS), and resonant impulse stimulated Raman spectroscopy (RISRS).\cite{RN103, Dhar1994, RN345, Cho2000, RN491}
Here we explore the less-utilized hyper-Raman transition in four wave mixing CMDS.

@@

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a$ and $\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$. 
While most hyper-Raman experiments are performed using degenerate input pulses ($\omega_a = \omega_b$), it is possible to perform the technique with non-degenerate input pulses ($\omega_a \neq \omega_b$). \cite{Denisov1986, Kozich2007}
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through a two-photon inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been observed in several CMDS studies (sometimes under the moniker of SIVE).\cite{Zilian1994, RN350, RN351, RN352, RN353, Chen1998, RN362, RN418, Hanninen2018, Wang2021, Bonn2024, McDonnell2024}

@@

Nonlinear spectroscopic probes of electronic-vibrational coupling are not limited to the Raman and hyper-Raman based techniques diagrammed in \autoref{fig:comparisonwmel}.
Independently developed by the Fleming and Khalil groups, respectively, 2D Electronic-Vibrational (2D-EV) and 2D Vibrational-Electronic (2D-VE) spectroscopy are $\chi^{(3)}$ methods based upon pump-probe and absorption type pathways which also investigate e-v coupling.\cite{Oliver2014, Courtney2015, Courtney2015_1}
The Rao group has also developed 2D-EV sum frequency generation (2D-EVSFG), a $\chi^{(4)}$ analogue of 2D-EV, to probe interfacial e-v coupling. \cite{Deng2021}  
These methods have been used to understand vibronic coupling in a variety of molecular systems and, albeit with different selection rules, serve as 2D pump-probe type analogues to the infrared-hyper-Raman spectroscopies discussed here.

When the hyper-Raman excitation in HDFG or HSFG is resonant or near-resonant with electronic states, the brightness of the vibrational feature depends on the vibronic structure of the electronic state.
The electronic spectrum probed in HDFG or HSFG can inform on processes that control ultrafast electronic relaxation in molecular and biological systems in ways similar to resonance Raman.\cite{Cho2001, Bredenbeck2015, Gaynor2018, Arsenault2021}
Since the selection rules of hyper-Raman scattering differ from Raman scattering, the hyper-Raman excitation spectra give a unique alternative to Raman scattering to understand vibrational spectra, electronic structure and vibronic coupling in molecular systems. \cite{Olson2018}

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
Coherent spectroscopies are often described using the density matrix formalism in the interaction picture, where an n$^{\text{th}}$ order density matrix $\rho^{(n)}$ is propagated using the Liouville-von Neumann equation:
\begin{equation}\label{LVNE}
	\dot{\rho}^{(n)} = \frac{1}{i \hbar} [H + V(t), \rho^{(n-1)}] + \dot{\rho}^{(n)}_R,
\end{equation}
where $H$ is the (static) system Hamiltonian, $V(t)$ is the field-matter interaction and $\rho_R$ is the relaxation density matrix. \cite{RN455}
In systems where interactions with surroundings are relevant, expressions for $\dot{\rho}_R$ may be quite complicated due to intricate phenomena such as system-bath interactions and solvent-solute coupling, among others. \cite{Sung2001}
In the limit where input field pulse widths are much longer than electronic dephasing times and the individual molecule limit is approached, it becomes acceptable to write a time-independent relaxation matrix $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$, i.e. exponential relaxation, where $\Gamma$ is the system's dephasing rate. \cite{RN455, Sue1986, Sung2001}
While exponential relaxation is usually an inappropriate description of most physical systems involving solvent-solute interactions,\cite{Sue1986, Yan1988, Li1994, Myers1997} our goal is to introduce simple, quantitative models to understand HDFG response.
Thus, we write $\dot{\rho}^{(n)}_R= \Gamma \rho^{(n)}$.
The exponential relaxation limit has proven useful in investigating the properties of 2D-EV, 2D-VE, and doubly resonant SFG using pulses much longer than electronic dephasing times. \cite{Raschke2002, Gaynor2017}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.\cite{Cho2008}
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or obscure coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=\textwidth]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
		The virtual states fluctuate in energy dependent upon the energy of the input fields.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is vibrationally resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned from an electronic state. \cite{McDonnell2024}
Note that in the case where $\omega_1$ interacts first but $\omega_2$ is resonant, the nonlinear output generated nearly vanishes as the first resonant pulse arrives after a non-resonant pulse.

@@

Other potential interferants in HDFG spectroscopy are direct and sequential cascades of second-order vDFG and vSFG processes.\cite{RN297, Cho2000_Cascade, RN301}
While unimportant in achiral isotropic systems,\cite{Belkin2000} second-order cascades may become important in media where SFG and DFG are allowed, e.g. chiral media, interfaces or non-centrosymmetric media. 
Nevertheless, in the discussion that follows for the rest of this manuscript, we will assume isotropic media, and therefore negligible second-order cascades.

It is useful to expose relationships between nonlinear output, transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit ($\dot{\rho}^{(n)} = 0 \ \forall n$). \cite{Moad2005}
Under the electric dipole approximation, the $I^\text{th}$ Cartesian component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation) \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \varepsilon_0 \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(- \omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the $IJKL$ element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}\label{eq:nfgamma}
	\chi^{(3)}_{IJKL} = NF \langle \gamma_{ijkl} \rangle_{IJKL}.
\end{equation}
For clarity, we define the input electric fields as plane waves, where (suppressing the spatial dependence for simplicity)
\begin{equation}
	\vec{E}\left(\omega \right) = \frac{1}{2} \left(\vec{E}_0 \exp\left(i \omega t\right) +\text{c.c.}\right) .
\end{equation}
In \autoref{eq:nfgamma}, $N$ is a number density, $F=\prod_j  \frac{1}{3} \left({\left(n(\omega_j)\right)^2 + 1} \right)$ is the Lorentz local field factor, where $n$ is the (frequency dependent) refractive index and $\gamma_{ijkl}$ is the third-order polarizability (i.e., second hyperpolarizability). \cite{Bedeaux1973}
The brackets indicate an orientational average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.
The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}

By propagating \autoref{LVNE} in the steady state limit,\cite{Prior1984} the HDFG hyperpolarizability is 
\begin{equation}\label{sivegamma}
		\gamma_{ijkl}^{vg} =	- \sum_{e,m} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{ve}_{i} \mu^{em}_{j} \mu^{mg}_{k} \mu^{gv}_{l} }{\Delta_{ev}^{3+2-1} \Delta_{mv}^{2-1}\Delta_{gv}^{-1}}  \rho^{(0)}_{gg}
\end{equation}
where: $\mu^{ab}_{j}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl}^{-x+y+z} = \omega_{kl} - (-\omega_{x}+\omega_{y} + \omega_{z}) - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the $\rho_{kl}$ dephasing, and $\rho^{(0)}_{gg}$ is the ground state population.
Here, $\ket{e}$ and $\ket{m}$ are virtual states, and thus necessarily summed over in \autoref{sivegamma}.
The pulse identifiers on the $\Delta_{kl}$ terms will be suppressed henceforth for legibility.
$D$, the Maker-Terhune degeneracy factor, accounts for permutation symmetry of the excitation fields.\cite{RN134} 
For a HDFG experiment using two (three) distinct, non-degenerate input fields, $D = 3 (6)$.

@@

It is useful to investigate the selection rules without vibronic resonance to understand how HDFG can be used as an upconverted infrared spectroscopy.
We first investigate the HDFG selection rules through a Placzek type formalism.
When $\omega_2+\omega_3$ is significantly detuned from resonance and the vibronic nature of the virtual states is ignored, the Placzek approximation can be invoked to study selection rules in terms of the hyper-Raman hyperpolarizability, $\beta$. \cite{Placzek1934, Long1970}
Summing over the virtual states $\ket{e}$ and $\ket{m}$ forms the hyper-Raman hyperpolarizability so that\cite{Long1970} 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{vg}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho^{(0)}_{gg}.
\end{equation}
Since all infrared active transitions are hyper-Raman active, HDFG is allowed for any infrared active transition. \cite{Andrews1978}
This selection rule is generally valid for any HDFG or HSFG process when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

The general selection rule can be simplified by considering the excitation of an infrared active normal mode.  
Assume $\ket{v}$ is the $v^{\text{th}}$ normal mode.
We Taylor expand the dipole and first hyperpolarizability operators to first order in the normal mode coordinate $Q_v$ about equilibrium:
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_v}\right)_0 Q_v 
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_v}\right)_0 Q_v
	\end{equation}
\end{subequations}
Taking $\ket{g} = \ket{0}, \ket{v} = \ket{1}$, substituting into \autoref{sivebeta} and using typical Harmonic Oscillator selection rules, gives the HDFG hyperpolarizability to $\order{Q_v}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{8D \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{vg}_{ijk}}{\partial Q_v}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_v}}\right)_0  \rho^{(0)}_{gg},
\end{equation}
where we have used $\mel{v}{\mu_{l,0}}{g} = 0 = \mel{1}{\beta_{ijk,0}}{0}$, and $\mel{1}{Q_v}{0} = \sqrt{\frac{\hbar}{2\omega_{vg}}}$.
Here, $\omega_{vg}$ is the characteristic $\ket{0} \rightarrow \ket{1}$ frequency.\cite{RN459}
Since \autoref{SIVEselection} is non-zero in the harmonic oscillator limit, HDFG output is allowed for infrared active normal modes in the harmonic limit.

\subsection{Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the general source of HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to Albrecht's treatment of Raman spectroscopy (i.e., sum-over-states formalism).\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(\textbf{Q})) \otimes \ket{b}$ for electronic states $\{|a(\textbf{Q}))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where \textbf{Q}, the system normal modes, is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on \textbf{Q} for simplicity.
It is worthwhile noting that the time-dependent Raman formalism developed by Heller \textit{et al.} can be applied to hyper-Raman scattering. \cite{Lee1979, Silverstein2012}
However, since our goal is to expose the Quantum Mechanical aspects of HDFG spectroscopy, we choose to employ the typical sum-over-states formalism.
